---
title: "Pilot - 0 - Preprocessing"
author: "René Michel"
date: "August 2020"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: united
---


Each analysis step will produce a markdown html documentation. Code chunks (except the specifications) will be hidden to keep the output readable. Thus, you can either knit the whole document using the "knit" button to run the whole script at once and directly see the output, or use the "play" buttons in the right upper corner of each code chunk to run the analysis step by step within RStudio.

```{r clear environment, include = FALSE}
# clear environment
rm(list = ls())
cat("\014")
```

```{r installations, include = FALSE}


# tested with 

# R 3.6.1 (you can check your r version with R.Version()$version.string)
# win10-x64
# RStudio 1.2.1335

# installs devtools if not found on machine, needed to install packages from github
if(!("devtools" %in% installed.packages()[,"Package"])) install.packages('devtools')
library('devtools')
# will install required dependencies for RAP analysis
devtools::install_github('remichel/envRAP') 
# initialize all required packages
rmTools::libraries(ggplot2, reshape2, pracma, Rmisc, circular, scales, R.matlab, dplyr, stats, lattice, car, rmTools, knitr, kableExtra) 
```

## Preprocessing Specifications

You can find all specifications for the applied preprocessing below.

```{r specifications}
subject_list        = c(1:9,12:16)
session_list        = c(1)
n_soas              = 3
do_setseed          = T    # set to T if you want to reproduce the findings
do_createdirs       = T    # if T, script checks for directories and creates them if necessary
do_save             = T    # T if you want to save files. might overwrite existing datafiles
do_mm               = T    # export matlab files for mixture model analysis?
```

```{r paths, include = FALSE}
study_path          = 'Z:/René/RATES/OSF_public/Pilot/'
data_path           = paste0(study_path,'Logfiles/')
plot_path           = paste0(study_path,'Plots/')
preprocessed_path   = paste0(study_path,'Preprocessed/')
if(do_createdirs) create_dirs(preprocessed_path, plot_path) # checks for paths and creates them if missing
```

```{r load data, include = FALSE}

# create helper vars
n_sub     = length(subject_list)
n_session = length(session_list)

# create dataset list
datasets  = as.data.frame(matrix(NA,n_sub,n_session))
for(iSub in 1:n_sub){
  for(iSession in 1:n_session){
    datasets[iSub, iSession] = paste(subject_list[iSub],"_",session_list[iSession], sep= "")
  }
}

# load & merge datasets
for (iSub in 1:n_sub){
  for(iSession in 1:n_session){
    if (iSub == 1 & iSession == 1) {
      data   = read.table(paste0(data_path,datasets[iSub,iSession],".txt"), sep="\t",h=T) 
    }else{
      if (file.exists(paste(data_path,datasets[iSub,iSession],".txt",sep=""))){
        data = rbind(data, read.table(paste0(data_path,datasets[iSub,iSession],".txt"), sep="\t",h=T))
      }
    }
  }
}
```


```{r filter data, include = FALSE}
# Exclude invalid trials (to slow responses & broken fixations), practice trials and first practice session

# mode = 0 (practice)
# mode = 1 (quest / familiarization phase, only session 1)
# mode = 2 (test trials)

raw       = data # keep raw dataset
data      = subset(data,data$mode > 1 & data$abort == "VALID")

```


```{r prepare dataset, include = FALSE}

# align SOAs
# logfiles contain exact screen timing estimates from PTB, therefore they may vary slightly across sessions
# set verbose or difference_plot to T to check for deviations after rounding
data$soa = align_soas(data$soa,n_soa = 3, verbose = F, difference_plot = F)
data$soa = round(data$soa)

# create / recode variables
data$target_position  = factor(data$target_position,levels = c("left","right"))
data$cue_position     = factor(data$cue_position,levels = c("left","right"))
data$validity         = as.factor(ifelse(data$cue_position == data$target_position, "valid", "invalid"))
data$rtms             = data$rt*1000

# compute error values
data$error   = as.numeric(data$target_degree)-as.numeric(data$reported_degree) # calculate raw error [-360, 360]
data$error   = ifelse(data$error > 180, 360-data$error,
                      ifelse(data$error < -180, -360-data$error, 
                             data$error)) # recode errors > 180 / <-180 to get error values in [-180, 180]
data$abserr           = abs(data$error) # recode to absolute values
data$correct_90 = ifelse(data$abserr < 90, 1, 0) # dichotomize absolute errors into hits/misses to check whether QUEST was successful
```

## Sanity Checks

The following tables serve to double-check all relevant cells, e.g. if any value is missing. 

```{r sanity checks, echo = F}

# check distribution of trials per cell
knitr::kable(xtabs(~data$subject+data$soa)) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>% 
  add_header_above(c("Subject", "SOA" = 3)) %>%
  scroll_box(width = "100%")
knitr::kable(xtabs(~data$subject+data$validity)) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>% 
  add_header_above(c("Subject", "Validity" = 2)) %>%
  scroll_box(width = "100%")

knitr::kable(xtabs(~data$subject[data$validity == "valid"]+data$soa[data$validity == "valid"])) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>% 
  add_header_above(c("Subject", "SOA" = 3)) %>%
  add_header_above(c("valid" = 4)) %>%
  scroll_box(width = "100%")

knitr::kable(xtabs(~data$subject[data$validity == "invalid"]+data$soa[data$validity == "invalid"])) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>% 
  add_header_above(c("Subject", "SOA" = 3)) %>%
  add_header_above(c("invalid" = 4)) %>%
  scroll_box(width = "100%")
```

## Performance Checks

Performance per subject. For the prestudy, we have not preregistered exlcusion criteria. Still, each subject would have been within the main study criteria for accuracy >.60 or <.80.

```{r performance checks, echo = F}

# check performance per session per participant
performance = aggregate(correct_90 ~ subject,data,mean)

knitr::kable(xtabs(performance$correct_90 ~ performance$subject)) %>% 
  kable_styling(bootstrap_options = c("striped", "hover")) %>% 
  add_header_above(c("Subject", "Session" = 1))  %>% 
  add_header_above(c("", "Performance (+/- 90° = hit)" = 1))  %>%
  scroll_box(width = "100%")

```

```{r r export, include = FALSE}
# export preprocessed data for further analyses and plotting
if(do_save)  save(raw, file   = paste0(preprocessed_path,"raw.Rdata"))
if(do_save)  save(data, file   = paste0(preprocessed_path,"preprocessed.Rdata"))
if(do_save)  save(performance, file = paste0(preprocessed_path,"performance.Rdata"))

```


```{r matlab export, include = FALSE}

# export as matlab files for mixture modelling

if(do_mm){
  
  full_data = data # store full dataset 
  
  for (iSub in subject_list){

    
    # subset to single subject
    data = full_data[full_data$subject == iSub,]
    
    # sort dataset by validity & SOA
    data = arrange(data,validity, soa)
    
      
    # create empty matrix for matlab data
    memfit  = as.data.frame(matrix(NA,max(xtabs(~data$validity+data$soa)),length(unique(data$validity))*length(unique(data$soa))))
    # write error values in matrix memfit columnwise for each SOA*validity combination
    for (j in 1:length(unique(data$validity))){
      for (i in 1:length(unique(data$soa))){
        k = (j-1)*length(unique(data$soa))+i # save error values in kth column with current error vector
        # save subset of current SOA/val combination in kth column
        memfit[,k]= subset(data, data$validity == unique(data$validity)[j] &  data$soa == unique(data$soa)[i])$error 
        # name column with current SOA*validity combination
        colnames(memfit)[k] = paste0(unique(data$validity)[j],'_',unique(data$soa)[i]) 
      }
    }
    
    # save colnames in separate names variable
    names = colnames(memfit)
    
    # save to files
    if(do_save) writeMat(paste0(preprocessed_path,iSub,"_mm_data.mat"), d = memfit)
    if(do_save) save(names, file=paste0(preprocessed_path,iSub,"_mm_names.R"))
    
  }
        
        
        
}
```


