---
title: "Pilot 2- Absolute Error & MM"
author: "René Michel"
date: "August 2020"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: united
---

Each analysis step will produce a markdown html documentation. Code chunks (except the specifications) will be hidden to keep the output readable. Thus, you can either knit the whole document using the "knit" button to run the whole script at once and directly see the output, or use the "play" buttons in the right upper corner of each code chunk to run the analysis step by step within RStudio.

```{r clear environment, include = FALSE}
# clear environment
rm(list = ls())
cat("\014")
```

```{r installations, include = FALSE}


# tested with 

# R 3.6.1 (you can check your r version with R.Version()$version.string)
# win10-x64
# RStudio 1.2.1335

# installs devtools if not found on machine, needed to install packages from github
if(!("devtools" %in% installed.packages()[,"Package"])) install.packages('devtools')
library('devtools')
# will install required dependencies for RAP analysis
devtools::install_github('remichel/envRAP') 
# initialize all required packages
rmTools::libraries(ggplot2, reshape2, pracma, Rmisc, circular, scales, R.matlab, dplyr, stats, lattice, car, rmTools, knitr, kableExtra) 
```

## Specifications

You can find all specifications for the applied preprocessing below.

```{r specifications}
subject_list        = c(1:9,12:16)
n_soa               = 3
do_setseed          = T    # set to T if you want to reproduce the findings
do_createdirs       = T    # if T, script checks for directories and creates them if necessary
do_save             = T    # T if you want to save files. might overwrite existing datafiles
```

```{r paths, include = FALSE}
study_path          = 'Z:/René/RATES/OSF_public/Pilot/'
mm_out_path         = paste0(study_path,'MM_Out/')
plot_path           = paste0(study_path,'Plots/')
preprocessed_path   = paste0(study_path,'Preprocessed/')
if(do_createdirs) create_dirs(plot_path) # checks for paths and creates them if missing
```

```{r helperfuns, include = F}
help_fun_path       = 'C:/Users/rmich/Buschlab/sortRAPcode/helper_funs/' 
helper_functions = paste0(help_fun_path, list.files(help_fun_path))
sapply(helper_functions, source)
```

```{r load data, include = FALSE}


# load MM data
MMfileEnd     = "_mm.txt"
NamesFileName = '_mm_names.R'


# Load data 
for (iSub in subject_list){
    if(iSub == min(subject_list)){
      mm = read.table(paste0(mm_out_path,iSub,MMfileEnd), h=T)
      load(paste0(preprocessed_path,iSub,NamesFileName))
      conds = as.data.frame(t(matrix(unlist(strsplit(names,'_')),2,2*n_soa)))
      subj = rep(iSub,length(mm[,1]))
      mm = cbind(subj,conds,mm)
    }else{
      tmp = read.table(paste0(mm_out_path,iSub,MMfileEnd), h=T)
      load(paste0(preprocessed_path,iSub,NamesFileName))
      conds = as.data.frame(t(matrix(unlist(strsplit(names,'_')),2,2*n_soa)))
      subj = rep(iSub,length(tmp[,1]))
      tmp = cbind(subj,conds,tmp)
      mm = rbind(mm,tmp)
    }
    if(iSub == max(subject_list)){
      colnames(mm)[2:3] = c('validity','soa')
      mm$permNum = 0
      rm(tmp, subj, conds)
    }
}

# load preprocessed data

pre = loadRData(paste0(preprocessed_path,'preprocessed.Rdata'))

```

# MM Tests

```{r compute differences, include = F}

# melt mm
long_mm = melt(mm, id.vars = c("subj", "soa", "validity"), measure.vars = c("g", "sd"))
# cast mm
cast_mm = dcast(long_mm, subj+soa+variable ~ validity, value.var = "value")
# compute diff
cast_mm$diff = cast_mm$valid-cast_mm$invalid


```

## MM Shapiro tests

```{r shapiros, echo = F, comment = ""}

# normality tests
knitr::kable(shapiroby(cast_mm$diff[cast_mm$variable == "g"], cast_mm$soa[cast_mm$variable == "g"])) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>% 
  add_header_above(c("MM g: Shapiro tests" = 3))  %>%
  scroll_box(width = "100%")
  
knitr::kable(shapiroby(cast_mm$diff[cast_mm$variable == "sd"], cast_mm$soa[cast_mm$variable == "sd"])) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>% 
  add_header_above(c("MM sd: Shapiro tests" = 3))  %>%
  scroll_box(width = "100%")
```

## MM tests 

```{r ttests, echo = F, comment = ""}
# ttests
gtests = tapply(cast_mm$diff[cast_mm$variable == "g"], cast_mm$soa[cast_mm$variable == "g"], t.test)
names(gtests) = c("g: 125 ms", "g: 158 ms", "g: 192 ms")
gtests

sdtests = tapply(cast_mm$diff[cast_mm$variable == "sd"], cast_mm$soa[cast_mm$variable == "sd"], t.test)
names(sdtests) = c("sd: 125 ms", "sd: 158 ms", "sd: 192 ms")
sdtests

```

## Wilcox tests

```{r wtests, echo = F, comment = ""}
# wilcox test for not normally distributed soa
wilcox.test(cast_mm$diff[cast_mm$variable == "sd" & cast_mm$soa == 125])
```


# Plots

```{r plots, echo = F, message = F, warnings = F, out.width = '100%'}

theme_set(theme_classic())

# average g/sd/abserr
g = summarySEwithin(mm, measurevar = "g", betweenvars = NULL, withinvars = c("validity","soa"), idvar = "subj", conf.interval = 0.95)
g$soa = fac2num(g$soa)

mm$mmsd = mm$sd
sd = summarySEwithin(mm, measurevar = "mmsd", betweenvars = NULL, withinvars = c("validity","soa"), idvar = "subj", conf.interval = 0.95)
sd$soa = fac2num(sd$soa)

mae = summarySEwithin(pre, measurevar = "abserr", betweenvars = NULL, withinvars = c("validity","soa"), idvar = "subject", conf.interval = 0.95)
mae$soa = fac2num(mae$soa)

# merge datasets
colnames(g)[4]    = "value"
colnames(sd)[4]   = "value"
colnames(mae)[4]  = "value"

g$var   = "sd"
sd$var  = "g"
mae$var = "mae"

merged = rbind(mae, g, sd)
merged$var = factor(merged$var, levels = c("mae","g","sd"))

# plot

pre_summary = ggplot(merged, aes(x=soa, y=value, group = validity))+
  geom_errorbar(aes(ymin = value-se, ymax = value+se, col = validity), size = 1, width = 10, alpha = .7)+
  geom_line(aes(col = validity), size = 3, alpha =1)+
  geom_point(aes(col = validity), fill = "white", shape = 21, stroke = 2,size = 6) + 
  scale_x_continuous(breaks = round(unique(merged$soa)), labels = round(unique(merged$soa)))+
  scale_color_manual(values=c("sienna2", "dodgerblue2"),breaks = levels(merged$validity))+
  xlab('time in ms')+
  ylab(NULL)+
  facet_wrap_custom(~var, 
                    scales = 'free_y', 
                    ncol = 3, 
                    strip.position = "left", 
                    labeller = as_labeller(c(mae = "Mean Absolute Error", g = "sd", sd = "g")),
                    scale_overrides = list(scale_override(1, scale_y_continuous(expand = c(.2,.2))),
                                           scale_override(2, scale_y_continuous(expand = c(0.03, 0.03)))))

pre_summary

```




```{r save plot, include = F}
theme_set(theme_classic())

pre_summary+ 
  theme(panel.border = element_rect(color = "black", fill = NA, size = 2.2),
        panel.spacing = unit(.8, "cm"),axis.line = element_line(colour = 'black', size = 2.2),
        panel.margin = unit(2,"cm"),
        strip.background = element_blank(),
        strip.placement = "outside",
        strip.text = element_text(colour = "black", size = 30, face = 'bold'),
        axis.ticks = element_line(colour = "black", size = 2.2),
        axis.ticks.length = unit(.4, "cm"),
        axis.title.x = element_text(size = 40,face = 'bold', margin = margin(t = 40, r = 0, b = 0, l = 0)),
        axis.text.y = element_text(size = 22,face = 'bold', margin = margin(t = , r = 12, b = 0, l = 0), color = 'black'),
        axis.text.x = element_text(size = 22,face = 'bold', margin = margin(t = 12, r = 0, b = 0, l = 0), color = 'black'),
        legend.text = element_text(size = 28,face = 'bold'),
        legend.title = element_blank(),
        legend.background = element_rect(color = "black", size = 1.2),
        legend.key.size = unit(18,"point"),
        legend.spacing.x = unit(.5,"cm"))

ggsave(filename = paste0(plot_path,'pilot_overview.png'), device = "png", height = 25, width = 60, units = "cm")
ggsave(filename = paste0(plot_path,'pilot_overview.pdf'), device = "pdf", height = 25, width = 60, units = "cm")



```
