---
title: "Main - 1 -Mean Absolute Error"
author: "René Michel"
date: "September 2020"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: united
---

Each analysis step will produce a markdown html documentation. Code chunks will be hidden to keep the output readable. Thus, you can either knit the whole document using the "knit" button to run the whole script at once and directly see the output, or use the "play" buttons in the right upper corner of each code chunk to run the analysis step by step within RStudio.

```{r clear environment, include = FALSE}
# clear environment
rm(list = ls())
cat("\014")
```

```{r installations, include = FALSE}


# tested with 

# R 3.6.1 (you can check your r version with R.Version()$version.string)
# win10-x64
# RStudio 1.2.1335

# installs devtools if not found on machine, needed to install packages from github
if(!("devtools" %in% installed.packages()[,"Package"])) install.packages('devtools')
library('devtools')
# will install required dependencies for RAP analysis
devtools::install_github('remichel/envRAP') 
# initialize all required packages
rmTools::libraries(ggplot2, reshape2, pracma, Rmisc, circular, scales, R.matlab, dplyr, stats, lattice, car, rmTools, knitr, kableExtra) 
```

## Specifications

```{r specifications}
do_createdirs       = T    # if T, script checks for directories and creates them if necessary
do_save             = T    # T if you want to save files. might overwrite existing datafiles
```


```{r paths, include = FALSE}
study_path          = 'D:/OSF/Main/'
preprocessed_path   = paste0(study_path,'Preprocessed/')
plot_path           = paste0(study_path,'Plots/')
if(do_createdirs) create_dirs(preprocessed_path, plot_path) # checks for paths and creates them if missing
```

```{r load data, include = FALSE}
main = loadRData(paste0(preprocessed_path,'preprocessed.Rdata'))
```

```{r compute mean absolute error, include = FALSE}

mae = summarySEwithin(main, measurevar = "abserr", betweenvars = NULL,
                           withinvars = c("validity","soa"), idvar = "subject",
                           conf.interval = 0.95)

mae$soa = round(fac2num(mae$soa))
```

## Plot

```{r plot, echo = F, out.width = '100%'}

mae_plot = ggplot(mae,aes(x = soa, y = abserr, group = validity))+
  geom_errorbar(aes(ymin = abserr-se, ymax = abserr+se, col = validity), size = 1, width = 15, alpha = .7)+
  geom_line(aes(col = validity), size = 3, alpha =1)+
  geom_point(aes(col = validity), fill = "white", shape = 21, stroke = 2,size = 6) + 
  xlab('SOA in ms')+
  ylab('Mean Absolute Error')+
  scale_x_continuous(breaks = round(unique(mae$soa))[seq(1,length(mae$soa),2)], 
                     labels = round(unique(mae$soa))[seq(1,length(mae$soa),2)])+
  scale_color_manual(values=c("sienna2", "dodgerblue2"),breaks = rev(levels(mae$validity)))+
  theme_classic()

mae_plot
```
  



```{r save, include = F}

# use theme commands to make a nice pdf / png 
mae_plot+
  theme(strip.text = element_text(colour = "black", size = 18, face = 'bold'),
        strip.text.x = element_text(size = 30),
        strip.background = element_rect(color = 'black', size =2.2),
        panel.border = element_rect(color = "black", fill = NA, size = 2.2),
        axis.line = element_line(colour = 'black', size = 2.2),
        axis.ticks = element_line(colour = "black", size = 2.2),
        axis.ticks.length = unit(.4, "cm"),
        axis.title.y = element_text(size = 40, face = 'bold',margin = margin(t = 0, r = 30, b = 0, l = 0)),
        axis.title.x = element_text(size = 40,face = 'bold', margin = margin(t = 40, r = 0, b = 0, l = 0)),
        axis.text.y = element_text(size = 30,face = 'bold', margin = margin(t = , r = 12, b = 0, l = 0), color = 'black'),
        axis.text.x = element_text(size = 28,face = 'bold', margin = margin(t = 12, r = 0, b = 0, l = 0), color = 'black'),
        legend.text = element_text(size = 28,face = 'bold'),
        legend.title = element_blank(),
        legend.key.size = unit(18,"point"),
        legend.key.width = unit(3, 'cm'), 
        legend.position = c("NONE"),
        legend.spacing.x = unit(.5,"cm"))

if(do_save) ggsave(filename = paste0(plot_path,'main_mae_timecourse.png'), device = "png", height = 25, width = 60, units = "cm")
if(do_save) ggsave(filename = paste0(plot_path,'main_mae_timecourse.pdf'), device = "pdf", height = 25, width = 60, units = "cm")
```