---
title: "Main - 6.1 - FFT Analysis Phase"
author: "René Michel"
date: "February 2021"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: united
---


Each analysis step will produce a markdown html documentation. Code chunks will be hidden to keep the output readable. Thus, you can either knit the whole document using the "knit" button to run the whole script at once and directly see the output, or use the "play" buttons in the right upper corner of each code chunk to run the analysis step by step within RStudio.

```{r clear environment, include = FALSE}
# clear environment
rm(list = ls())
cat("\014")
```

```{r installations, include = FALSE}

# tested with 

# R 3.6.1 (you can check your r version with R.Version()$version.string)
# win10-x64
# RStudio 1.2.1335

# installs devtools if not found on machine, needed to install packages from github
if(!("devtools" %in% installed.packages()[,"Package"])) install.packages('devtools')
library('devtools')
# will install required dependencies for RAP analysis
devtools::install_github('remichel/envRAP') 
# initialize all required packages
rmTools::libraries(ggplot2, reshape2, pracma, Rmisc, circular, scales, R.matlab, dplyr, stats, lattice, car, rmTools, knitr, kableExtra) 
```

## Specifications

### General

```{r specifications}
subject_list        = c(1:10,12:15)
n_soa               = 20 
do_createdirs       = T    # if T, script checks for directories and creates them if necessary
do_save             = T    # T if you want to save files. might overwrite existing datafiles
```

### FFT Specifications

```{r fft specs}
# FFT specifications
sfreq               = 1000*1/(233.4-191.7) # in Hz
desired_res         = 1.2 # in Hz
desired_out         = 'amp'
outphase            = T 
padding             = 'none'
detrend             = 'linear'
window              = 'none'
scaling             = 'z'
```

```{r paths, include = FALSE}
study_path          = 'D:/OSF/Main/'
preprocessed_path   = paste0(study_path,'Preprocessed/') 
mm_out_path         = paste0(study_path,'MM_Out/') 
fft_out_path        = paste0(study_path,'FFT_Out/')  
if(do_createdirs) create_dirs(fft_out_path) # checks for paths and creates them if missing
```


```{r load data, include = F}

# in files
MMfileEnd           = "_mm.txt"
NamesFileName       = '_mm_names.R'
# out files
MMpermEnd           = "_mm_perm.txt"
Outname_Data        = "mm_data.Rdata"
Outname_Perm        = "mm_perm.Rdata"
Outname_All         = "mm_all.Rdata"
Outname_Phase_Wide  = "mm_phase_wide.Rdata"
Outname_Phase_Long  = "mm_phase_long.Rdata"

# load real data 
for (iSub in subject_list){
  if(iSub == min(subject_list)){
    data = read.table(paste0(mm_out_path,iSub,MMfileEnd), h=T)
    load(paste0(preprocessed_path,iSub,NamesFileName))
    conds = as.data.frame(t(matrix(unlist(strsplit(names,'_')),2,2*n_soa)))
    subj = rep(iSub,length(data[,1]))
    data = cbind(subj,conds,data)
  }else{
    tmp = read.table(paste0(mm_out_path,iSub,MMfileEnd), h=T)
    load(paste0(preprocessed_path,iSub,NamesFileName))
    conds = as.data.frame(t(matrix(unlist(strsplit(names,'_')),2,2*n_soa)))
    subj = rep(iSub,length(tmp[,1]))
    tmp = cbind(subj,conds,tmp)
    data = rbind(data,tmp)
  }
  if(iSub == max(subject_list)){
    colnames(data)[2:3] = c('validity','soa')
    rm(tmp, subj, conds)
  }
}


if(do_save) save(data,file = paste0(fft_out_path,Outname_Data))
```


```{r prepare for fft, include = F}

# bring into wide format
sd      = dcast(data, subj + validity ~ soa, value.var = "sd")
g       = dcast(data, subj + validity ~ soa, value.var = "g")

sd$var  = 'sd'
g$var   = 'g'
wide    = rbind(sd,g)
wide    = wide[moveme(names(wide),'var first')] # rearrange dataframe, bring id.vars to front of dataframe
  
```



```{r fft, include = F}

firstLine = 4 # indicate first line of data in dataframe
  
# Extract phase
phase = dsa_fft(wide[firstLine:(dim(wide)[2])], 
              time = 'col', 
              detrend = detrend, 
              window = window, 
              scaling = scaling , 
              padding = padding, 
              desired_res = desired_res, 
              sfreq = sfreq, 
              desired_out = desired_out, 
              single_sided = T, 
              out = 'arg', # define phase as output
              alpha = alpha, 
              verbose = T )

# Concatenate with identifier vars again and save wideformat datasets
phase = cbind(wide[1:(firstLine-1)], phase)
if(do_save) save(phase,file = paste0(fft_out_path, Outname_Phase_Wide))

# Reshape to longformat for ggplot
phase_long = melt(phase, id.vars = c('subj','validity','var'), variable.name = 'freq', value.name = 'phase')
if(do_save) save(phase_long,file = paste0(fft_out_path, Outname_Phase_Long))

```

## Rayleigh Test Results

```{r Rayleigh tests, echo = F, comment = ""}

# round freqs
phase_long$freq = round(fac2num(phase_long$freq),1)

# specify which conditions shall be tested

testlist = list(freq = c(4.8, 9.6, 10.8), val = c("invalid", "invalid", "valid"), m = c("g", "sd", "sd"))

for(i in 1:length(testlist[[1]])){
  data = phase_long[phase_long$validity == testlist$val[i] &
                    phase_long$freq == testlist$freq[i] &
                    phase_long$var == testlist$m[i],]
  
  data$phase = circular(data$phase, unit = "radians")

  fprintf(paste(testlist$m[i], testlist$freq[i], testlist$val[i], "\n"))
  print(rayleigh.test(data$phase))
}

```