---
title: "Main - 3 - Validity Analysis"
author: "René Michel"
date: "August 2020"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: united
---

Each analysis step will produce a markdown html documentation. Code chunks will be hidden to keep the output readable. Thus, you can either knit the whole document using the "knit" button to run the whole script at once and directly see the output, or use the "play" buttons in the right upper corner of each code chunk to run the analysis step by step within RStudio.

```{r clear environment, include = FALSE}
# clear environment
rm(list = ls())
cat("\014")
```

```{r installations, include = FALSE}

# tested with 

# R 3.6.1 (you can check your r version with R.Version()$version.string)
# win10-x64
# RStudio 1.2.1335

# installs devtools if not found on machine, needed to install packages from github
if(!("devtools" %in% installed.packages()[,"Package"])) install.packages('devtools')
library('devtools')
# will install required dependencies for RAP analysis
devtools::install_github('remichel/envRAP') 
# initialize all required packages
rmTools::libraries(ggplot2, reshape2, pracma, Rmisc, circular, scales, R.matlab, dplyr, stats, lattice, car, rmTools, knitr, kableExtra) 
```

## Specifications

```{r specifications}
subject_list        = c(1:10,12:15)
n_soa               = 20 
do_createdirs       = T    # if T, script checks for directories and creates them if necessary
do_save             = T    # T if you want to save files. might overwrite existing datafiles
```


```{r paths, include = FALSE}
study_path          = 'C:/Data/OSF_public/Main/'
preprocessed_path   = paste0(study_path,'Preprocessed/')
mm_out_path         = paste0(study_path,'MM_Out/') 
plot_path           = paste0(study_path,'Plots/')
if(do_createdirs) create_dirs(plot_path) # checks for paths and creates them if missing
```

```{r load mm data, include = FALSE}

MMfileEnd     = "_mm.txt"
NamesFileName = '_mm_names.R'


# Load data 
for (iSub in subject_list){
    if(iSub == min(subject_list)){
      data = read.table(paste0(mm_out_path,iSub,MMfileEnd), h=T)
      load(paste0(preprocessed_path,iSub,NamesFileName))
      conds = as.data.frame(t(matrix(unlist(strsplit(names,'_')),2,2*n_soa)))
      subj = rep(iSub,length(data[,1]))
      data = cbind(subj,conds,data)
    }else{
      tmp = read.table(paste0(mm_out_path,iSub,MMfileEnd), h=T)
      load(paste0(preprocessed_path,iSub,NamesFileName))
      conds = as.data.frame(t(matrix(unlist(strsplit(names,'_')),2,2*n_soa)))
      subj = rep(iSub,length(tmp[,1]))
      tmp = cbind(subj,conds,tmp)
      data = rbind(data,tmp)
    }
    if(iSub == max(subject_list)){
      colnames(data)[2:3] = c('validity','soa')
      data$permNum = 0
      rm(tmp, subj, conds)
    }
}

```


```{r filter data, include = F}

# get the first SOA only
solo = data[data$soa == unique(data$soa)[1],]

# get the first 3 SOAs
three = data[as.numeric(data$soa) <= as.numeric(unique(data$soa))[3],]
# average first three SOAs
three_g = summarySE(three, measurevar = "g", groupvars = c("subj", "validity"))
names(three)[which(names(three) == "sd")] = "mmsd"
three_sd = summarySE(three, measurevar = "mmsd", groupvars = c("subj", "validity"))

```


```{r compute differences, include = F}

# compute differences val - inval

diff_solo_g = na.omit(solo %>% arrange(validity, subj) %>% group_by(subj) %>% mutate(diff = c(diff(g), NA)))
diff_solo_sd = na.omit(solo %>% arrange(validity, subj) %>% group_by(subj) %>% mutate(diff = c(diff(sd), NA)))

diff_three_g = na.omit(three_g %>% arrange(validity, subj) %>% group_by(subj) %>% mutate(diff = c(diff(g), NA)))
diff_three_sd = na.omit(three_sd %>% arrange(validity, subj) %>% group_by(subj) %>% mutate(diff = c(diff(mmsd), NA)))

```

## Test Results

The tested difference for all tests is valid-invalid

## First SOA g
```{r test solo g, echo = F, comment = ""}

shapiro.test(diff_solo_g$diff)
t.test(diff_solo_g$diff, alternative = "less")

```

## First SOA sd
```{r test solo sd, echo = F, comment = ""}

shapiro.test(diff_solo_sd$diff)
t.test(diff_solo_sd$diff, alternative = "less")

```

## First 3 SOA g
```{r test three g, echo = F, comment = ""}

shapiro.test(diff_three_g$diff)
t.test(diff_three_g$diff, alternative = "less")

```

## First 3 SOA sd
```{r test three sd, echo = F, comment = ""}

shapiro.test(diff_three_sd$diff)
t.test(diff_three_sd$diff, alternative = "less")

```


## Violin Plot

```{r, echo = F, message = FALSE, out.width = '100%'}

# average g across subjects
gavg = summaryRMwithin(three_g, 
                         measurevar = "g", 
                         withinvars = c("validity"), 
                         idvar = c("subj"), 
                         doNorm = F, # to leave the absolute values untouched, we just want the adjusted SE estimates for withinvars
                         conf.interval = 0.95)

violinplot = ggplot(NULL)+
        geom_violin(data = three_g, aes(x = validity, y = g, group = validity, fill = validity))+
        geom_point(data = three_g, aes(x = validity, y = g, group = subj), size = 2.5)+
        geom_line(data = three_g, aes(x = validity, y = g, group = subj), size = 1, alpha = .3)+
        geom_point(data = gavg, aes(x = validity, y = g) ,col = "black", size = 6)+
        geom_errorbar(data = gavg, aes(x = validity, ymin = g-se, ymax = g+se), size = 1.2, width = .2)+
        geom_line(data = gavg, aes(x = validity, y = g, group = 1) ,col = "black",size = 3)+
        geom_smooth(data = gavg, aes(x = validity, y = g) , method = lm, se = T,size = 3)+
        scale_fill_manual(values=c("sienna2", "dodgerblue2"),breaks = levels(gavg$validity))+
        labs(y = "g")+
        theme_classic()

violinplot

```

```{r save plot, include = F}

theme_RM = set_theme_RM()

violinplot + 
  coord_cartesian(ylim = c(.39, .83))+
  theme_RM +
  theme(legend.position = "none",
        axis.title.x = element_text(size = 40,face = 'bold', margin = margin(t = 40, r = 40, b = 40, l = 40)),
        axis.title.y = element_text(size = 40,face = 'bold', margin = margin(t = 40, r = 40, b = 0, l = 40)),
        axis.text.y = element_text(size = 22,face = 'bold', margin = margin(t = , r = 12, b = 0, l = 0), color = 'black'),
        axis.text.x = element_text(size = 22,face = 'bold', margin = margin(t = 12, r = 0, b = 0, l = 0), color = 'black'))

if(do_save) ggsave(paste0(plot_path, "main_mm_g_validity_effect.png"), device = "png", height = 22, width = 22, units = "cm")
if(do_save) ggsave(paste0(plot_path, "main_mm_g_validity_effect.pdf"), device = "pdf", height = 22, width = 22, units = "cm")
```

