---
title: "Main - 5.1 - FFT Analysis Amplitude"
author: "René Michel"
date: "August 2020"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: united
---

Each analysis step will produce a markdown html documentation. Code chunks will be hidden to keep the output readable. Thus, you can either knit the whole document using the "knit" button to run the whole script at once and directly see the output, or use the "play" buttons in the right upper corner of each code chunk to run the analysis step by step within RStudio.

```{r clear environment, include = FALSE}
# clear environment
rm(list = ls())
cat("\014")
```

```{r installations, include = FALSE}

# tested with 

# R 3.6.1 (you can check your r version with R.Version()$version.string)
# win10-x64
# RStudio 1.2.1335

# installs devtools if not found on machine, needed to install packages from github
if(!("devtools" %in% installed.packages()[,"Package"])) install.packages('devtools')
library('devtools')
# will install required dependencies for RAP analysis
devtools::install_github('remichel/envRAP') 
# initialize all required packages
rmTools::libraries(ggplot2, reshape2, pracma, Rmisc, circular, scales, R.matlab, dplyr, stats, lattice, car, rmTools, knitr, kableExtra) 
```

## Specifications

### General

```{r specifications}
subject_list        = c(1:10,12:15)
n_soa               = 20 
do_createdirs       = T     # if T, script checks for directories and creates them if necessary
do_save             = T     # T if you want to save files. might overwrite existing datafiles
do_skip_load        = T     # load merged mm file directly
do_skip_fft         = T     # loaf fft output directly
```

### FFT Specifications

```{r fft specs}
# FFT specifications
sfreq               = 1000*1/(233.4-191.7) # in Hz
desired_res         = 1.2 # in Hz
desired_out         = 'amp'
outphase            = T 
padding             = 'none'
detrend             = 'linear'
window              = 'none'
scaling             = 'z'
```

### Permutation Test Specifications

```{r test specs}
alphalevel          = .95 # family-wise alphalevel
mcc_ntests          = 10 # number of tests to correct for (Bonferroni)
```

```{r paths, include = FALSE}
study_path          = 'Z:/René/RATES/OSF_public/Main/'
preprocessed_path   = paste0(study_path,'Preprocessed/') 
mm_out_path         = paste0(study_path,'MM_Out/') 
fft_out_path        = paste0(study_path,'FFT_Out/')  
plot_path           = paste0(study_path,'Plots/')
if(do_createdirs) create_dirs(fft_out_path, plot_path) # checks for paths and creates them if missing
```


```{r load data, include = F}

# in files
MMfileEnd           = "_mm.txt"
NamesFileName       = '_mm_names.R'
# out files
MMpermEnd           = "_mm_perm.txt"
Outname_Data        = "mm_data.Rdata"
Outname_Perm        = "mm_perm.Rdata"
Outname_All         = "mm_all.Rdata"
Outname_Amp_Wide    = "mm_amp_wide.Rdata"
Outname_Amp_Long    = "mm_amp_long.Rdata"
Outname_GA_SD       = "mm_ga_sd.Rdata"


# check whether merged file shall be loaded
if(do_skip_load){
  loadStatus = !file.exists(paste0(mm_out_path, Outname_All))
  if(!loadStatus) load(paste0(mm_out_path, Outname_All))
}else{
  loadStatus = T
}


if(loadStatus){
  # load real data 
  for (iSub in subject_list){
    if(iSub == min(subject_list)){
      data = read.table(paste0(mm_out_path,iSub,MMfileEnd), h=T)
      load(paste0(preprocessed_path,iSub,NamesFileName))
      conds = as.data.frame(t(matrix(unlist(strsplit(names,'_')),2,2*n_soa)))
      subj = rep(iSub,length(data[,1]))
      data = cbind(subj,conds,data)
    }else{
      tmp = read.table(paste0(mm_out_path,iSub,MMfileEnd), h=T)
      load(paste0(preprocessed_path,iSub,NamesFileName))
      conds = as.data.frame(t(matrix(unlist(strsplit(names,'_')),2,2*n_soa)))
      subj = rep(iSub,length(tmp[,1]))
      tmp = cbind(subj,conds,tmp)
      data = rbind(data,tmp)
    }
    if(iSub == max(subject_list)){
      colnames(data)[2:3] = c('validity','soa')
      data$permNum = 0
      rm(tmp, subj, conds)
    }
  }



  # load permuted data
  for (iSub in subject_list){
    if(iSub == min(subject_list)){
      perm = read.table(paste0(mm_out_path,iSub,MMpermEnd), h=T)
      n_perm = length(perm[,1])/(n_soa*2)
      load(paste0(preprocessed_path,iSub,NamesFileName))
      conds = as.data.frame(t(matrix(unlist(strsplit(names,'_')),2,2*n_soa)))
      subj = rep(iSub,length(perm[,1]))
      permNum = sort(rep(c(1:n_perm),2*n_soa))
      cond = rep_dataframe(conds,n_perm,by = 'rbind')
      perm = cbind(subj,cond,perm,permNum)
    }else{
      tmp = read.table(paste0(mm_out_path,iSub,MMpermEnd), h=T)
      n_perm = length(tmp[,1])/(n_soa*2)
      load(paste0(preprocessed_path,iSub,NamesFileName))
      conds = as.data.frame(t(matrix(unlist(strsplit(names,'_')),2,2*n_soa)))
      subj = rep(iSub,length(tmp[,1]))
      permNum = sort(rep(c(1:n_perm),2*n_soa))
      cond = rep_dataframe(conds,n_perm,by = 'rbind')
      tmp = cbind(subj,cond,tmp,permNum)
      # concatenate
      perm = rbind(perm,tmp)
    }
    if(iSub == max(subject_list)){
      colnames(perm)[2:3] = c('validity','soa')
      rm(tmp, subj, cond, conds, permNum)
    }
    
  }
  
  # save dataframes
  if(do_save) save(perm,file = paste0(mm_out_path,Outname_Perm))
  if(do_save) save(data,file = paste0(mm_out_path,Outname_Data))
  
  # concatenat
  data = rbind(data,perm)
  rm(perm)
  if(do_save) save(data,file = paste0(mm_out_path,Outname_All))
}

```


```{r prepare for fft, include = F}

# check whether FFT shall be performed or not
if(do_skip_fft){
  doFFT = !(file.exists(paste0(fft_out_path,Outname_Amp_Long)) & file.exists(paste0(fft_out_path,Outname_GA_SD)))
  if(!doFFT) load(paste0(fft_out_path,Outname_Amp_Long))
  if(!doFFT) load(paste0(fft_out_path,Outname_GA_SD))
}else{
  doFFT = T
}

if(doFFT){
  
  # rename variable for Rmisc functions
  data$mmsd = data$sd
  
  # Average across participants
  averageTC_sd = summarySE(data, measurevar = 'mmsd', groupvars = c('permNum','validity', 'soa'), conf.interval = .95)
  averageTC_g = summarySE(data, measurevar = 'g', groupvars = c('permNum','validity', 'soa'), conf.interval = .95)
  
    
  # Reshape and concatenate
  sd      = dcast(averageTC_sd, permNum + validity ~ soa, value.var = "mmsd")
  sd$var  = 'sd'
  g       = dcast(averageTC_g, permNum + validity ~ soa, value.var = "g")
  g$var   = 'g'
  averageWide    = rbind(sd,g)
  averageWide    = averageWide[moveme(names(averageWide),'var first')] # rearrange dataframe, bring id.vars to front of dataframe
  if(do_save) save(averageTC_sd,file = paste0(fft_out_path,Outname_GA_SD))
}

```

```{r fft, include = F}

# check whether FFT shall be performed or not
if(do_skip_fft){
  doFFT = !file.exists(paste0(fft_out_path,Outname_Amp_Long))
  if(!doFFT) load(paste0(fft_out_path,Outname_Amp_Long))
  if(!doFFT) load(paste0(fft_out_path,Outname_GA_SD))
}else{
  doFFT = T
}

if(doFFT){
  
  firstLine = 4 # indicate first line of data in dataframe
    
  # Extract Amplitude
  amp = dsa_fft(averageWide[firstLine:dim(averageWide)[2]], 
                time = 'col', 
                detrend = detrend, 
                window = window, 
                scaling = scaling , 
                padding = padding, 
                desired_res = desired_res, 
                sfreq = sfreq, 
                desired_out = desired_out, 
                single_sided = T, 
                out = 'mod', # define amplitude as output
                alpha = alpha, 
                verbose = T )
  
  # Concatenate with identifier vars again and save wideformat
  amp = cbind(averageWide[1:(firstLine-1)], amp)
  if(do_save) save(amp,file = paste0(fft_out_path,Outname_Amp_Wide))
  
  # Reshape to longformat for ggplot
  amp_long = melt(amp, id.vars = c('permNum','validity','var'), variable.name = 'freq', value.name = 'amplitude')
  if(do_save) save(amp_long,file = paste0(fft_out_path,Outname_Amp_Long))
}

```

```{r perm test, include = F}

perms_long    = subset(amp_long,amp_long$permNum > 0)
amp_long      = subset(amp_long,amp_long$permNum == 0)


# reshape permutations to wideformat
perms_wide = dcast(perms_long, validity + var + freq ~ permNum, value.var = "amplitude")
# sort datasets in same way to concatenate them later on
amp_long    = arrange(amp_long, validity, var, freq)
perms_wide  = arrange(perms_wide, validity, var, freq)

# identify threshold amplitudes for .95 quantile
firstLine = 4 # indicate first line of data in dataframe
amp_long$thres = apply(perms_wide[,firstLine:dim(perms_wide)[2]],1, quantile, probs = alphalevel)

# compute p values for observed amplitudes
for(i in 1:dim(perms_wide)[1]){
  amp_long$p[i] = length(which(perms_wide[i,firstLine:dim(perms_wide)[2]] > amp_long$amplitude[i]))/length(perms_wide[1,firstLine:dim(perms_wide)[2]])
}

# identify threshold amplitudes for Bonferroni corrected alphalevel
amp_long$thres_bonf = apply(perms_wide[,firstLine:dim(perms_wide)[2]],1,quantile, probs = 1-(1-alphalevel)/mcc_ntests)
  
```

## Permutation test results 

```{r perm test results, echo = F}

results = amp_long[!(colnames(amp_long) %in% c("permNum", "amplitude", "thres", "thres_bonf"))]
results$freq = round(fac2num(results$freq),1)
results$sig = ifelse(results$p < (1-alphalevel), "*", "")
results$sig_after_bonf = ifelse(results$p < (1-alphalevel)/mcc_ntests, "*", "")

knitr::kable(results) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>% 
  scroll_box(width = "100%")
```

## Amplitude Spectrum

```{r plot spectrum, echo = F, out.width = '100%'}

# bring into longformat
thres = melt(amp_long, id.vars = c('validity','var','freq', 'amplitude'), variable.name = 'method', value.name = 'thres')
# exclude unwanted vars
thres = thres[thres$method != "p" &  thres$method != "permNum",]

# prepare for plot
thres$freq = fac2num(thres$freq)
  
xlabels = rep('',length(unique(thres$freq)))
xlabels[seq(1,length(xlabels),2)] = round(unique(thres$freq)[seq(1,length(xlabels),2)],1)

  
  
  
amp_plot = ggplot(data = thres, aes(x = freq, y = amplitude, group = validity))+
  facet_wrap(~var, ncol = 2)+
  geom_ribbon(data = thres[thres$method == "thres_bonf",], aes(ymin = 0, ymax = thres,fill = validity), alpha =.2)+ # ribbon for bonferroni
  geom_ribbon(data = thres[thres$method == 'thres',], aes(ymin = 0, ymax = thres,fill = validity), alpha =.3)+ # ribbon for.95 quantile
  geom_line(aes(color = validity), size = 12)+
  scale_x_continuous(breaks = unique(thres$freq), labels = xlabels, expand = c(0,0))+
  scale_y_continuous(limits = c(0,13), breaks = seq(0,12,2),expand = c(0,0))+
  scale_color_manual(values=c("sienna2", "dodgerblue2"),breaks = levels(thres$validity))+  #rev(levels(thres$validity))
  scale_fill_manual(values=c("sienna2", "dodgerblue2"),breaks = levels(thres$validity))+
  xlab('Frequency')+
  ylab('Amplitude (AU)')+
  theme_light()
  
  
# add significance asterics
if(any(amp_long$amplitude > amp_long$thres_bonf)){
  amp_long$sig = NA
  amp_long$sig[which(amp_long$amplitude > amp_long$thres_bonf)] = '*'
  asterics = na.omit(amp_long)
  asterics$amplitude = asterics$amplitude*1.05
  asterics$freq = fac2num(asterics$freq)
  amp_plot = amp_plot  + geom_text(data = asterics, aes(label = sig), size = 22)
  amp_plot
}else{
  amp_plot
}


```

```{r save amp plot, include = F}

amp_plot + 
  theme(strip.text = element_text(colour = "black", size = 18, face = 'bold'),
        strip.text.x = element_text(size = 30),
        strip.background = element_rect(color = 'black', size =2.2),
        panel.border = element_rect(color = "black", fill = NA, size = 2.2),
        panel.spacing = unit(.8, "cm"),
        axis.line = element_line(colour = 'black', size = 2.2),
        axis.ticks = element_line(colour = "black", size = 2.2),
        axis.ticks.length = unit(.4, "cm"),
        axis.title.y = element_text(size = 40, face = 'bold',margin = margin(t = 0, r = 30, b = 0, l = 0)),
        axis.title.x = element_text(size = 40,face = 'bold', margin = margin(t = 40, r = 0, b = 0, l = 0)),
        axis.text.y = element_text(size = 30,face = 'bold', margin = margin(t = , r = 12, b = 0, l = 0), color = 'black'),
        axis.text.x = element_text(size = 28,face = 'bold', margin = margin(t = 12, r = 0, b = 0, l = 0), color = 'black'),
        legend.key.width = unit(3, 'cm'), 
        legend.text = element_text(size = 28,face = 'bold'),
        legend.title = element_blank(),
        legend.key.size = unit(18,"point"),
        legend.position = c(.15,.9),
        legend.background = element_rect(color = 'black', size = 1.5, fill = "white"),
        legend.spacing.x = unit(.5,"cm"))
  

if(do_save) ggsave(filename = paste0(plot_path,'main_mm_spectrum.png'), device = "png", height = 32, width = 40, units = "cm")
if(do_save) ggsave(filename = paste0(plot_path,'main_mm_spectrum.pdf'), device = "pdf", height = 32, width = 40, units = "cm")

```

## Permutation test (for significant peak)

```{r plot perm test, echo = F, out.width = '100%'}

# which condition is significant?
signif = which(amp_long$amplitude > amp_long$thres_bonf)

# Extract significant combination
sig_freq      = amp_long$freq[signif]
sig_validity  = amp_long$validity[signif]
sig_var       = amp_long$var[signif]

# Subset to significant condition and concatenate
sig_line  = amp_long[amp_long$var == sig_var & amp_long$validity == sig_validity & amp_long$freq == sig_freq,]
distr     = perms_long[perms_long$var == sig_var & perms_long$validity == sig_validity & perms_long$freq == sig_freq,]
plotdata  = rbind(sig_line[,1:dim(distr)[2]], distr)

# Plot
permplot = ggplot(data = plotdata[plotdata$permNum > 0,], aes(x = amplitude))+
  geom_histogram(binwidth = .15, fill = "sienna2", alpha = .5)+
  geom_vline(xintercept = sig_line$thres, size = 6, color = 'grey10', linetype = 'dashed')+
  geom_vline(xintercept = sig_line$thres_bonf, size = 6, color = 'grey70', linetype = 'dashed')+
  geom_vline(xintercept = sig_line$amplitude, size = 6, color = "sienna2", linetype = "solid")+
  scale_x_continuous(breaks = seq(0,12,2),expand = c(0,0))+
  scale_y_continuous(expand = c(0,0))+
  xlab("Amplitude (AU)")+
  ylab("Number of Permutations")+
  theme_classic()

permplot
```

```{r save perm plot, include = F}
  
permplot+
  theme(panel.border = element_rect(color = "black", fill = NA, size = 3),
        axis.line = element_line(colour = 'black', size = 3),
        axis.ticks = element_line(colour = "black", size = 3),
        axis.ticks.length = unit(.6, "cm"),
        axis.title.y = element_text(size = 54, face = 'bold',margin = margin(t = 0, r = 20, b = 0, l = 0)),
        axis.title.x = element_text(size = 54,face = 'bold', margin = margin(t = 30, r = 0, b = 0, l = 0)),
        axis.text.y = element_text(size = 44,face = 'bold', margin = margin(t = , r = 12, b = 0, l = 0), color = 'black'),
        axis.text.x = element_text(size = 44,face = 'bold', margin = margin(t = 12, r = 0, b = 0, l = 0), color = 'black'),
        legend.title = element_blank(),
        legend.key.size = unit(18,"point"),
        legend.position = "top",
        legend.direction = "horizontal",
        legend.spacing.x = unit(.5,"cm"))

if(do_save) ggsave(filename = paste0(plot_path,'main_mm_sd_permtest.png'), device = "png", height = 40, width = 40, units = "cm")
if(do_save) ggsave(filename = paste0(plot_path,'main_mm_sd_permtest.pdf'), device = "pdf", height = 40, width = 40, units = "cm")

```

## Sinusoidal fit plot

```{r sinusoid fit, echo = F, out.width = '100%'}

# time vectors
t         = fac2num(unique(data$soa)) # discrete for observed data
time      =  c(min(t):max(t)) # continuous for predictions

# detrend observed values
obs_dat   = detrend(averageTC_sd$mmsd[averageTC_sd$permNum == 0 & averageTC_sd$validity == sig_validity])
obs_se    = averageTC_sd$se[averageTC_sd$permNum == 0 & averageTC_sd$validity == sig_validity] # we need SEs for plot again

# concatenate & reshape detrended data for plotting
plotdat   = as.data.frame(cbind(t, obs_dat, obs_se))
colnames(plotdat) = c("t", "obs_dat", "obs_se")
plotdat   = melt(plotdat, id.vars = c('t', 'obs_se'), variable.name = 'type', value.name = "sd")

# fit sinusoid with fixed frequency
freqms    = fac2num(sig_freq)/1000 # define fixed frequency
fit       = lm(obs_dat ~ sin(2*pi*freqms*t) + cos(2*pi*freqms*t)) # fitting
pred_dat  = fit$coefficients[1] + fit$coefficients[2]*sin(2*pi*freqms*time) + fit$coefficients[3]*cos(2*pi*freqms*time) # predictions

# concatenate & reshape predicted data for plotting
plotpred  = as.data.frame(cbind(time, pred_dat))
colnames(plotpred) = c("time", "pred_dat")
plotpred   = melt(plotpred, id.vars = c('time'), variable.name = 'type', value.name = "sd")


# define labels
labels = rep("",length(t))
labels[seq(1,length(t),2)] = round(unique(t))[seq(1,length(t),2)]

osc_plot = ggplot(plotdat[plotdat$type == "obs_dat",], aes(x = t, y = sd))+
  geom_errorbar(aes(ymin = sd-obs_se, ymax = sd+obs_se), color = "sienna2",size = 1.5, width = 6, alpha = 1) + 
  geom_line(aes(color = "sienna2"), size = 5, alpha =1)+
  geom_point(size = 12, color = "sienna2", stroke = 2) + 
  geom_line(data = plotpred[plotpred$type == "pred_dat",], aes(x = time, y = sd, color = "black"), size = 7, alpha = 1)+
  scale_y_continuous()+ 
  scale_x_continuous(breaks = t, labels = labels)+
  scale_color_manual(values=c("black","sienna2"),labels = c("fitted 9.6 Hz oscillation","observed data"))+
  xlab('time in ms')+
  ylab('sd (detrended)')+
  theme_classic()

osc_plot

```


```{r save osc plot, include = F}
osc_plot + theme(axis.line = element_line(colour = 'black', size = 2.2),
        axis.ticks = element_line(colour = "black", size = 2.2),
        axis.ticks.length = unit(.4, "cm"),
        panel.border = element_rect(color = "black", fill = NA, size = 2.4),
        axis.title.y = element_text(size = 40, face = 'bold',margin = margin(t = 0, r = 30, b = 0, l = 0)),
        axis.title.x = element_text(size = 40,face = 'bold', margin = margin(t = 40, r = 0, b = 0, l = 0)),
        axis.text.y = element_text(size = 30,face = 'bold', margin = margin(t = , r = 12, b = 0, l = 0), color = 'black'),
        axis.text.x = element_text(size = 28,face = 'bold', margin = margin(t = 12, r = 0, b = 0, l = 0), color = 'black'),
        legend.text = element_text(size = 28,face = 'bold'),
        legend.title = element_blank(),
        legend.background = element_rect(color = "black", size = 1.2),
        legend.key.size = unit(18,"point"),
        legend.position = c(.17,.11),
        legend.spacing.x = unit(.5,"cm"))

if(do_save) ggsave(filename = paste0(plot_path,'main_mm_sd_sinusoidal_fit.png'), device = "png", height = 20, width = 50, units = "cm")
if(do_save) ggsave(filename = paste0(plot_path,'main_mm_sd_sinusoidal_fit.pdf'), device = "pdf", height = 20, width = 50, units = "cm")
```
  