---
title: "Rev 1.4 FFT Single Participant Plot Acrossval"
author: "René Michel"
date: "Februar 2021"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: united
---


Each analysis step will produce a markdown html documentation. Code chunks will be hidden to keep the output readable. Thus, you can either knit the whole document using the "knit" button to run the whole script at once and directly see the output, or use the "play" buttons in the right upper corner of each code chunk to run the analysis step by step within RStudio.

```{r clear environment, include = FALSE}
# clear environment
rm(list = ls())
cat("\014")
```

```{r installations, include = FALSE}

# tested with 

# R 3.6.1 (you can check your r version with R.Version()$version.string)
# win10-x64
# RStudio 1.2.1335

# installs devtools if not found on machine, needed to install packages from github
if(!("devtools" %in% installed.packages()[,"Package"])) install.packages('devtools')
library('devtools')
# will install required dependencies for RAP analysis
devtools::install_github('remichel/envRAP') 
# initialize all required packages
rmTools::libraries(ggplot2, reshape2, pracma, Rmisc, circular, scales, R.matlab, dplyr, stats, lattice, car, rmTools, knitr, kableExtra) 
```

## Specifications

### General

```{r specifications}
subject_list        = c(1:10,12:15)
n_soa               = 20 
do_createdirs       = T     # if T, script checks for directories and creates them if necessary
do_save             = T     # T if you want to save files. might overwrite existing datafiles
```

### FFT Specifications

```{r fft specs}
# FFT specifications
sfreq               = 1000*1/(233.4-191.7) # in Hz
desired_res         = 1.2 # in Hz
desired_out         = 'amp'
outphase            = T 
padding             = 'none'
detrend             = 'linear'
window              = 'none'
scaling             = 'z'
```

```{r paths, include = FALSE}
study_path          = 'D:/OSF/Main/'
preprocessed_path   = paste0(study_path,'Preprocessed/') 
mm_out_path         = paste0(study_path,'MM_Out/') 
plot_path           = paste0(study_path,'Plots/')
if(do_createdirs) create_dirs(plot_path) # checks for paths and creates them if missing
```


```{r load data, include = F}

# in files
MMfileEnd           = "_mm_acrossval.txt"
NamesFileName       = '_mm_names_acrossval.R'



# load real data 
for (iSub in subject_list){
  if(iSub == min(subject_list)){
      data = read.table(paste0(mm_out_path,iSub,MMfileEnd), h=T)
      load(paste0(preprocessed_path,iSub,NamesFileName))
      conds = as.data.frame(t(matrix(unlist(strsplit(names,'_')),2,n_soa)))
      subj = rep(iSub,length(data[,1]))
      data = cbind(subj,conds,data)
  }else{
      tmp = read.table(paste0(mm_out_path,iSub,MMfileEnd), h=T)
      load(paste0(preprocessed_path,iSub,NamesFileName))
      conds = as.data.frame(t(matrix(unlist(strsplit(names,'_')),2,n_soa)))
      subj = rep(iSub,length(tmp[,1]))
      tmp = cbind(subj,conds,tmp)
      data = rbind(data,tmp)
  }
  if(iSub == max(subject_list)){
      colnames(data)[2:3] = c('validity','soa')
      data$permNum = 0
      rm(tmp, subj, conds)
  }
}
  
```


```{r prepare for fft, include = F}
data$mmsd = data$sd
sd      = dcast(data, permNum + validity + subj ~ soa, value.var = "mmsd")
sd$var = 'sd'
g      = dcast(data, permNum + validity + subj ~ soa, value.var = "g")
g$var = 'g'
Wide    = rbind(sd,g)
Wide    = Wide[moveme(names(Wide),'var first')] # rearrange dataframe, bring id.vars to front of dataframe

```


```{r extract single participant phase and amplitude, echo = F}

firstLine = 5 # indicate first line of data in dataframe

# compute fft and extract complex output
complex = dsa_fft(Wide[firstLine:dim(Wide)[2]], 
                time = 'col', 
                detrend = detrend, 
                window = window, 
                scaling = scaling , 
                padding = padding, 
                desired_res = desired_res, 
                out = "complex",
                sfreq = sfreq, 
                single_sided = T, 
                alpha = alpha, 
                verbose = T )

# merge output with id vars
complex = cbind(Wide[1:(firstLine-1)], complex)
# convert into long format
complex_long = melt(complex, id.vars = c('permNum','validity','var', 'subj'), variable.name = 'freq', value.name = 'complex')
# round freq
complex_long$freq = round(fac2num(complex_long$freq),1)
# extract amplitude and phase
complex_long$phase = Arg(complex_long$complex)
complex_long$amplitude = Mod(complex_long$complex)
# subset to relevant data for plot
complex_g = complex_long[complex_long$var == "g" & complex_long$freq == "4.8",]


```
## Single Participant Plot (G, across validity, 4.8Hz)


```{r plot single participants, echo = F, out.width = '100%'}
ggplot(complex_g, aes(x = phase, y = amplitude)) +
  geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf), fill = "grey80", alpha = .2)+
  geom_hline(yintercept = seq(0,10,2), color = "grey90", size = 0.5)+ # grid
  geom_segment(aes(x = phase, xend = phase,y = 0, yend = amplitude),color = "black", size =2, arrow = arrow(length = unit(0.5, "cm")))+
  coord_polar(start = pi/2, direction = -1)+
  scale_x_continuous(expand = c(0,0), limits = c(-pi,pi), breaks = seq(-pi,pi/2,pi/2), labels = c("180°", "270°", "0/360°", "90°"))+
  scale_y_continuous(expand = c(0,0), breaks = seq(0,10,2), limits = c(0,10))+
  ylab('Amplitude')+
  xlab('Phase Angle')+
  theme_minimal()+
  theme(strip.text.x = element_blank(),
        strip.text.y = element_blank(),
        strip.background = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_line(colour = "black", size = 1.2),
        axis.ticks.length = unit(.4, "cm"),
        axis.title.y = element_text(size = 40, face = 'bold',margin = margin(t = 0, r = 30, b = 0, l = 0)),
        axis.title.x = element_text(size = 40, face = 'bold',margin = margin(t = 0, r = 30, b = 0, l = 0)),
        axis.text.y = element_text(size = 18,face = 'bold', margin = margin(t = , r = 12, b = 0, l = 0), color = 'black'),
        axis.text.x = element_text(size = 18,face = 'bold', margin = margin(t = 12, r = 0, b = 0, l = 0), color = 'black'),
        legend.position = "none")

if (do_save) ggsave(filename = paste0(plot_path,'main_mm_fft_ssubj_g_acrossval_4Hz.pdf'), device = "pdf", height = 25, width = 25, units = "cm")

```
